---
title: "Zygosity & mutation class data prep"
author: "Maria Perry"
date: "2024-06-24"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(magrittr)
library(data.table)
library(dtplyr)

```

Input:

-   gxb_cohort_drivers

-   gxb_cohort_vus

-   gxb_cohort_combined

Output:

-   zyg_mut_class

    -   1 row per sample per gene containing:

        1.  biallelic label (monoallelic, biallelic, WT)
        2.  mutation class

# Load

```{r}

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

indir <- "data/input/"
outdir <- "data/parsed/"

gxb_cohort_drivers <- fread(paste0(indir, "gxb_cohort_drivers.csv"))
gxb_cohort_vus <- fread(paste0(indir, "gxb_cohort_vus.csv"))
gxb_cohort_combined <- fread(paste0(indir, "gxb_cohort_combined.csv"))

fga_tmb_blacklist <- fread(paste0(indir, "high_mutational_cna_burden_tumors.50k.txt"))
fga_tmb_blacklist <- fga_tmb_blacklist$x

```

# Mutation class

We will bind the driver and VUS tables together and remove duplicates. Only *mutations* are considered as driver or WT (including compound muts). Other forms of biallelic inactivation (homdel, etc.), fusions, and indeterminate zygosity calls are excluded (NA). All others are considered WT.

```{r}

drivers_vus <- rbindlist(list(drivers = gxb_cohort_drivers %>% lazy_dt() %>%
  # just annotate mut
  mutate(mutation_class = ifelse(grepl("Mut|compound", zygosity_call, ignore.case = T), "driver", NA_character_)) %>%
    as.data.table(),
  vus = gxb_cohort_vus %>% lazy_dt() %>%
  mutate(mutation_class = ifelse(grepl("Mut|compound", zygosity_call, ignore.case = T), "vus", NA_character_)) %>% as.data.table()), fill = T) %>% 
  lazy_dt() %>%
  distinct() %>%
  mutate(mutation_class = case_when(zygosity_call %in% c("Biallelic - Homdel", 
                                                         "Heterozygous (Fusion)",  "Indeterminate") ~ NA_character_,
                                    is.na(mutation_class) ~ "wt",
                                    TRUE ~ mutation_class)) %>%
  as.data.table()

drivers_vus <- drivers_vus %>%
  lazy_dt() %>%
  mutate(mutation_class_rank = as.numeric(factor(replace_na(mutation_class, "n/a"), levels = c("wt", "vus", "driver", "n/a"), ordered = T))) %>% # easier to take max with explicit NA level
  group_by(Tumor_Sample_Barcode, Hugo_Symbol) %>%
  filter(mutation_class_rank == max(mutation_class_rank)) %>% ungroup() %>%
  select(-mutation_class_rank) %>%
  as.data.table()

drivers_vus %>% xtabs(formula = ~ zygosity_call + mutation_class, addNA = T)

drivers_vus <- drivers_vus %>% lazy_dt() %>% select(Tumor_Sample_Barcode, Hugo_Symbol, mutation_class) %>% unique() %>% as.data.table()

```

# Biallelic label

*Only* for *KEAP1* in LUAD, we'll consider either drivers or VUS for the sake of zygosity based on initial analyses. For all other genes & cancer types, we use drivers only.

Biallelic label is categorized based on the following criteria: 

|                            |                              |
|----------------------------|------------------------------|
| **biallelic label**        | **zygosity_call**            |
| **biallelic**              | Biallelic-compound           |
| Biallelic-Homdel           |                              |
| Biallelic-Mut+fusion       |                              |
| Biallelic-Mut+LOH          |                              |
| **monoallelic**            | Gain-of-mutant-Mut+copygain  |
| Heterozygous(Mutation)     |                              |
| **wt**                     | Heterozygous-LOH(gene-level) |
| WT                         |                              |
| Amplification-(gene-level) |                              |
| Amplification-CNA          |                              |

## KEAP1 LUAD: driver or VUS

```{r}

data_samples_50k <- fread(paste0(indir, "data_clinical_sample.oncokb.txt.gz")) # use this for cancer type/oncotree code -- manually curated from 50k

gene_x_barcode_drivers_vus_keap1 <- gxb_cohort_drivers %>% 
  left_join(data_samples_50k %>% select(Tumor_Sample_Barcode = SAMPLE_ID, ONCOTREE_CODE)) %>%
  filter(Hugo_Symbol == "KEAP1" & ONCOTREE_CODE == "LUAD") %>% 
  mutate(mutation_class = ifelse(grepl("Mut", zygosity_call), "driver", NA)) %>%
  rbind(gxb_cohort_vus %>% 
          left_join(data_samples_50k %>% select(Tumor_Sample_Barcode = SAMPLE_ID, ONCOTREE_CODE)) %>% 
          filter(Hugo_Symbol == "KEAP1" & ONCOTREE_CODE == "LUAD")  %>% 
  mutate(mutation_class = ifelse(grepl("Mut", zygosity_call), "vus", NA))) %>%
  unique() %>%
  arrange(Tumor_Sample_Barcode, Hugo_Symbol, mutation_class) %>%
  mutate(biallelic_label = case_when(grepl("Biallelic", zygosity_call, ignore.case = T) ~ "biallelic",
                                     grepl("Mut", zygosity_call, ignore.case = T) ~ "monoallelic",
                                     zygosity_call %in% c("Indeterminate", "Heterozygous (Fusion)") ~ "n/a",
                                     TRUE ~ "wt"),
         biallelic_label = factor(biallelic_label, levels = c("n/a", "wt", "monoallelic", "biallelic"), ordered = T), 
         mutation_class = factor(replace_na(mutation_class, "n/a"), levels = c("n/a", "vus", "driver"), ordered = T)) %>% # easier to take max with explicit NA level 
  group_by(Tumor_Sample_Barcode, Hugo_Symbol) %>% filter(biallelic_label == max(biallelic_label)) %>%
  select(-c(is_oncogene, is_tsg, type)) %>%
  mutate(mutation_class = fct_na_level_to_value(mutation_class, "n/a"),
         biallelic_label = fct_na_level_to_value(biallelic_label, "n/a"))

gene_x_barcode_drivers_vus_keap1

gene_x_barcode_drivers_vus_keap1 %>% filter(Hugo_Symbol == "KEAP1") %>% xtabs(formula = ~ zygosity_call + biallelic_label, addNA = T)

```

## Drivers only (all other genes, types)

```{r}

gene_x_barcode_labeled <- gxb_cohort_drivers %>% lazy_dt() %>% 
  mutate(mutation_class = ifelse(grepl("Mut", zygosity_call), "driver", NA_character_)) %>% 
  mutate(biallelic = !is.na(zygosity_call) & grepl("Biallelic", zygosity_call, ignore.case = T),
         biallelic_label = case_when(biallelic == T ~ "biallelic",
                                     biallelic == F & grepl("Mut", zygosity_call, ignore.case = T) ~ "monoallelic",
                                     zygosity_call %in% c("Indeterminate", "Heterozygous (Fusion)") ~ NA_character_,
                                     TRUE ~ "wt"),
         biallelic_label = factor(biallelic_label, levels = c("wt", "monoallelic", "biallelic"), ordered = T)) %>%
  as.data.table()

gene_x_barcode_labeled %>% xtabs(formula = ~ zygosity_call + biallelic_label)

```

## Combined zygosity table

```{r}

# combine
gxb_zyg_labeled <- gene_x_barcode_drivers_vus_keap1 %>%
  rbind(gene_x_barcode_labeled %>% 
          left_join(data_samples_50k %>% select(Tumor_Sample_Barcode = SAMPLE_ID, ONCOTREE_CODE)) %>%
          filter(!(Hugo_Symbol == "KEAP1" & ONCOTREE_CODE == "LUAD")) %>% select(-c(is_oncogene, is_tsg, type))) 

gxb_zyg_labeled <- gxb_zyg_labeled %>%
  lazy_dt() %>%
  select(Tumor_Sample_Barcode, Hugo_Symbol, biallelic_label) %>% unique() %>% as.data.table()

```

# Combined zyg & mutation class data

```{r}

tsgs <- gxb_cohort_combined %>% lazy_dt() %>% filter(is_tsg == T & is_oncogene == F) %>% select(Hugo_Symbol) %>% unique() %>%
  as.data.table() %>% pull(Hugo_Symbol)

zyg_mut_class <- gxb_zyg_labeled %>%
  full_join(drivers_vus) %>%
  filter(Hugo_Symbol %in% tsgs) %>%
  filter(!Tumor_Sample_Barcode %in% fga_tmb_blacklist)
  

n_distinct(zyg_mut_class$Hugo_Symbol)

fwrite(zyg_mut_class, file = paste0(outdir, "zyg_mut_class.csv"))

```
