---
title: "KEAP1 OS and Chemo-IO PFS by mutation class and zygosity"
output: 
  html_document:
    toc: true
date: "2023-07-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(cowplot)
library(readr)
library(tools)
library(ggrepel)
library(RColorBrewer)
library(plotly)
library(svglite)
library(data.table)
library(readxl)
library(gridExtra)
library(colorspace)
library(broom)
library(binom)
library(rstatix)
library(ggpubr)
library(ggnewscale)
library(patchwork)
library(survminer)
library(survival)
library(ggupset)
library(magrittr)

```

# Load data / setup

```{r}

indir <- indir
outdir <- outdir
source("os_helper_functions.R")


alt_colors <- setNames(c("coral2", "peachpuff", "#7184A1", "#7184A1", 
                         "coral2", "peachpuff", "gray",
                         "skyblue3", "lightblue",  "lightblue", 
                         "gray", "gray", "gray50"), 
                       c("Driver", "VUS", "WT", "Wildtype",
                         "driver", "vus", "wt",
                         "biallelic", "het", "monoallelic", 
                         "wt", "N/A (WT)", "N/A (SV)"))

gene_name = "KEAP1"
os_inputs <- get(load(paste0(indir, "os_input_keap1.Robj")))
pfs_inputs <- get(load(paste0(indir, "pfs_input_keap1.Robj"))) %>% 
  mutate(driver_label = fct_recode(driver_label, WT = "wt"))

cox_models <- list()
km_models <- list()
cox_plots <- list()
km_plots <- list()

```


# OS

```{r}

os_covars <- c("msi_score", "level_1", "age_dx_days", "metastatic", "sex", "smoking_status", "cvr_tmb_score")

```

## Mutation class
### Multivariate (Cox PH model)
```{r}
model = coxph(as.formula(paste0("Surv(os_months, os_status) ~ mutation_class +", paste0(os_covars, collapse = " + "))) , data = os_inputs)
cox_models[[paste0(gene_name, "_driver")]] <- model
print(summary(model))
forest <- ggforest(model, main = paste0(gene_name, ": OS"), data = os_inputs)
print(forest)
cox_plots[[paste0(gene_name, "_driver")]] <- forest
```

### Kaplan-Meier curve

```{r}
surv = survfit(Surv(os_months, os_status) ~ mutation_class, data = os_inputs %>% mutate(mutation_class = fct_rev(mutation_class)))
names(surv$strata) <- gsub("mutation_class=", "", names(surv$strata))
km_models[[paste0(gene_name, "_driver")]] <- surv
gg <- ggsurvplot(surv, conf.int = F, pval = F, short.panel.labs = T, direction = -1, title = "", #paste0(gene_name, " (OS)"
                 legend.position = c(0.8, 0.8))
gg <- gg$plot + scale_color_manual(values = alt_colors, labels = c("Driver", "VUS", "Wildtype")) + 
  scale_x_continuous(breaks = seq(0, 120, 12*2)) +
  labs(color = "", y = "Overall survival probability", x = "Time (mo)") +
  theme(legend.position = c(0.8, 0.8), plot.title = element_text(hjust = 0.5))
km_plots[[paste0(gene_name, "_driver")]] <- gg
print(gg)
```

### Median OS & hazard ratios
```{r, echo=FALSE, results='asis', warning=F}
cat("#### OS")
cat("\n")
cat("##### Mutation class")
cat("\n")
cat("\n")
cat("VUS vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_driver, "VUS", "WT"),
       ", ", 
       pretty_HR_text(cox_models$KEAP1_driver, term = "mutation_classVUS"), "\n"))
cat("\n")
cat("Driver vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_driver, "Driver", "WT"),
       ", ", 
       pretty_HR_text(cox_models$KEAP1_driver, term = "mutation_classDriver"),  "\n"))
cat("\n")
cat("Driver vs VUS\n")
cat(pretty_survmedian_text(km_models$KEAP1_driver, "Driver", "VUS"), "\n")
```

## Zygosity
### Multivariate (Cox PH model)
```{r}
model = coxph(as.formula(paste0("Surv(os_months, os_status) ~ biallelic_label +", paste0(os_covars, collapse = " + "))) , data = os_inputs)
cox_models[[paste0(gene_name, "_zyg")]] <- model
#print(summary(model))
forest <- ggforest(model, main = paste0(gene_name, ": OS"), data = os_inputs)
cox_plots[[paste0(gene_name, "_zyg")]] <- forest
print(forest)
```

### Kaplan-Meier curve
```{r}
surv = survfit(Surv(os_months, os_status) ~ biallelic_label, data = os_inputs %>% mutate(biallelic_label = fct_rev(biallelic_label)))
names(surv$strata) <- gsub("biallelic_label=", "", names(surv$strata))
km_models[[paste0(gene_name, "_zyg")]] <- surv
gg <- ggsurvplot(surv, conf.int = F, pval = F, short.panel.labs = T, direction = -1, title = "", #paste0(gene_name, " (OS)"
                 legend.position = c(0.8, 0.8))
gg <- gg$plot + scale_color_manual(values = alt_colors, labels = c("Biallelic", "Monoallelic", "Wildtype")) + 
  scale_x_continuous(breaks = seq(0, 120, 12*2)) +
  labs(color = "", y = "Overall survival probability", x = "Time (mo)") +
  theme(legend.position = c(0.8, 0.8), plot.title = element_text(hjust = 0.5), plot.background = element_rect(fill = NA))
km_plots[[paste0(gene_name, "_zyg")]] <- gg
print(gg)
```

### Median OS & hazard ratios

```{r, echo=FALSE, results='asis', warning=F}
cat("#### OS")
cat("\n")
cat("##### Zyg")
cat("\n")
cat("\n")
cat("biallelic vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_zyg, "biallelic", "wt"),
       ", ",
       pretty_HR_text(cox_models$KEAP1_zyg, term = "biallelic_labelbiallelic")), "\n")
cat("\n")
cat("monoallelic vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_zyg, "monoallelic", "wt"),
       ", ", 
       pretty_HR_text(cox_models$KEAP1_zyg, term = "biallelic_labelmonoallelic")
), "\n")
cat("\n")
cat("biallelic vs monoallelic\n")
cat(pretty_survmedian_text(km_models$KEAP1_zyg, "biallelic", "monoallelic"), "\n")
```

# PFS
```{r}

pfs_covars <- c("impact_tmb", "sex", "pack_yrs", "pdl1_status", "dnlr", "albumin", "treatment_type_detail")
```

## Mutation class
### Multivariate (Cox PH model)
```{r}
model = coxph(as.formula(paste0("Surv(pfs_mo, progression) ~ driver_label +", paste0(pfs_covars, collapse = " + "))) , data = pfs_inputs)
cox_models[[paste0(gene_name, "_driver_pfs")]] <- model
print(summary(model))
forest <- ggforest(model, main = paste0(gene_name, ": PFS"), data = pfs_inputs)
print(forest)
cox_plots[[paste0(gene_name, "_driver_pfs")]] <- forest
```

### Kaplan-Meier curve
```{r}
surv = survfit(Surv(pfs_mo, progression) ~ driver_label, data = pfs_inputs %>% mutate(driver_label = fct_rev(driver_label)))
names(surv$strata) <- gsub("driver_label=", "", names(surv$strata))
km_models[[paste0(gene_name, "_driver_pfs")]] <- surv
gg <- ggsurvplot(surv, conf.int = F, pval = F, short.panel.labs = T, direction = -1, title = "", #paste0(gene_name, " (OS)"
                 legend.position = c(0.8, 0.8))
gg <- gg$plot + scale_color_manual(values = alt_colors, labels = (c("Driver", "VUS", "Wildtype"))) + 
  scale_x_continuous(breaks = seq(0, 120, 12*2)) +
  labs(color = "", y = "Progression-free survival probability", x = "Time (mo)") +
  theme(legend.position = c(0.8, 0.8), plot.title = element_text(hjust = 0.5))
print(gg)
km_plots[[paste0(gene_name, "_driver_pfs")]] <- gg
```

### Median PFS & hazard ratios
```{r, echo=FALSE, results='asis', warning=F}
cat("#### PFS")
cat("\n")
cat("##### Mutation class")
cat("\n")
cat("\n")
cat("VUS vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_driver_pfs, "vus", "WT"),
       ", ", 
       pretty_HR_text(cox_models$KEAP1_driver_pfs, term = "driver_labelvus"), "\n"))
cat("\n")

cat("Driver vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_driver_pfs, "driver", "WT"),
       ", ", 
       pretty_HR_text(cox_models$KEAP1_driver_pfs, term = "driver_labeldriver"),  "\n"))
cat("\n")
cat("Driver vs VUS\n")
cat(pretty_survmedian_text(km_models$KEAP1_driver_pfs, "driver", "vus"), "\n")
```

## Zygosity
### Multivariate (Cox PH model)
```{r}
model = coxph(as.formula(paste0("Surv(pfs_mo, progression) ~ biallelic_label +", paste0(pfs_covars, collapse = " + "))) , 
              data = pfs_inputs %>% filter(!biallelic_label %in% c("indeterminate", "N/A (SV)")) %>% mutate(biallelic_label = fct_drop(biallelic_label)))
cox_models[[paste0(gene_name, "_zyg_pfs")]] <- model
print(summary(model))
forest <- ggforest(model, main = paste0(gene_name, ": PFS"), data = pfs_inputs %>% filter(!biallelic_label %in% c("indeterminate", "N/A (SV)")) %>% mutate(biallelic_label = fct_drop(biallelic_label)))
print(forest)
cox_plots[[paste0(gene_name, "_zyg_pfs")]] <- forest
```


### Kaplan-Meier curve
```{r}
surv = survfit(Surv(pfs_mo, progression) ~ biallelic_label, data = pfs_inputs %>% filter(!biallelic_label %in% c("indeterminate", "N/A (SV)")) %>% mutate(biallelic_label = fct_drop(biallelic_label)) %>% mutate(biallelic_label = fct_rev(biallelic_label)))
names(surv$strata) <- gsub("biallelic_label=", "", names(surv$strata))
km_models[[paste0(gene_name, "_zyg_pfs")]] <- surv
gg <- ggsurvplot(surv, conf.int = F, pval = F, short.panel.labs = T, direction = -1, title = "", #paste0(gene_name, " (OS)"
                 legend.position = c(0.8, 0.8))
gg <- gg$plot + scale_color_manual(values = alt_colors, labels = c("Biallelic", "Monoallelic", "Wildtype")) + 
  scale_x_continuous(breaks = seq(0, 120, 12*2)) +
  labs(color = "", y = "Progression-free survival probability", x = "Time (mo)") +
  theme(legend.position = c(0.8, 0.8), plot.title = element_text(hjust = 0.5), plot.background = element_rect(fill = NA))
print(gg)
km_plots[[paste0(gene_name, "_zyg_pfs")]] <- gg
```

### Median PFS & hazard ratios
```{r, echo=FALSE, results='asis', warning=F}
cat("#### PFS")
cat("\n")
cat("##### Zyg")
cat("\n")
cat("\n")
cat("biallelic vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_zyg_pfs, "biallelic", "wt"),
       ", ",
       pretty_HR_text(cox_models$KEAP1_zyg_pfs, term = "biallelic_labelbiallelic")), "\n")
cat("\n")
cat("monoallelic vs WT\n")
cat(paste0(pretty_survmedian_text(km_models$KEAP1_zyg_pfs, "monoallelic", "wt"),
       ", ", 
       pretty_HR_text(cox_models$KEAP1_zyg_pfs, term = "biallelic_labelmonoallelic")
), "\n")
cat("\n")
cat("biallelic vs monoallelic\n")
cat(pretty_survmedian_text(km_models$KEAP1_zyg_pfs, "biallelic", "monoallelic"), "\n")


```

# Combined figures
```{r}
wrap_plots(km_plots)

```

```{r}
wrap_plots(cox_plots)
```

