---
title: "TSG OS"
output: html_document
date: "2023-07-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(cowplot)
library(readr)
library(tools)
library(ggrepel)
library(RColorBrewer)
library(plotly)
library(svglite)
library(data.table)
library(gridExtra)
library(rstatix)
library(ggpubr)
library(patchwork)
library(survminer)
library(survival)
library(magrittr)
library(gtable)
library(grid)
library(lubridate)
library(gtsummary)
set_gtsummary_theme(theme_gtsummary_compact())
library(dtplyr)
library(parallel)
library(survival)
library(survminer)
library(ggsurvfit)

```

# Setup

```{r}

indir <- "data/"
outdir <- "results/"

source("helper_functions_tsg.R")

```

# Load data

```{r}

zyg_mut_class <- fread(paste0(indir, "parsed/zyg_mut_class.csv"), na.strings = "")
data_os <- fread(paste0(indir, "input/data_os_pancan_revision_ALL.csv"), na.strings = "")

cohort_table_1 <- fread(paste0(indir, "input/Table_S1_clinDataFiltered.txt"))

```

```{r}

data_os_zyg <- data_os %>%
  lazy_dt() %>%
  inner_join(zyg_mut_class, by = c("sample_id" = "Tumor_Sample_Barcode")) %>%
  mutate(mutation_class_combined = ifelse(mutation_class %in% c("driver", "vus"), "altered", mutation_class),
         biallelic_label_combined = ifelse(biallelic_label %in% c("biallelic", "monoallelic"), "altered", biallelic_label)) %>%
  as.data.table()

fwrite(data_os_zyg, paste0(indir, "parsed/data_os_zyg.csv"), row.names = F)

```

# Function setup

## Gene/cancer type pairs for testing

Cancer types from Table 1 with at least 100 patients & 10 patients with 5 events per group

```{r}

pairs_to_test <- list()

covars_list <- c("disease_status", "sex", "age_at_dx", "oncokb_level_1", "fga", "tmb", "msi_score") 
required_fields <- c("os_status", "os_months", "oncotree_code", covars_list)

```

```{r}

# screen_input <- data_os_zyg %>% lazy_dt() %>%
#   select(sample_id, mutation_class, Hugo_Symbol, all_of(required_fields)) %>% drop_na() %>%
#   as.data.table()
# fwrite(screen_input, paste0(indir, "os_screen_input_with_covars.csv"))

```

### Mutation class

```{r}

pairs_to_test[["mutation_class"]] <- data_os_zyg %>% lazy_dt() %>%
  select(sample_id, mutation_class, Hugo_Symbol, all_of(required_fields)) %>% drop_na() %>% 
  filter(oncotree_code %in% cohort_table_1$ONCOTREE_CODE) %>%
  group_by(oncotree_code) %>% mutate(n_type = n_distinct(sample_id)) %>%
  group_by(Hugo_Symbol, oncotree_code, mutation_class) %>% add_count() %>% mutate(n_events = sum(os_status)) %>%
  group_by(Hugo_Symbol, oncotree_code) %>% filter(n_type >= 100 & n_distinct(mutation_class) == 3 & min(n) >= 10 & min(n_events) >= 5) %>% ungroup() %>%
  select(Hugo_Symbol, oncotree_code) %>% unique() %>% as.data.table()

n_distinct(pairs_to_test$mutation_class$Hugo_Symbol)
n_distinct(pairs_to_test$mutation_class$oncotree_code)
nrow(pairs_to_test$mutation_class)

  
```

### Zygosity (biallelic label)

```{r}

pairs_to_test[["biallelic_label"]] <- data_os_zyg %>% lazy_dt() %>%
  select(sample_id, biallelic_label, Hugo_Symbol, all_of(required_fields)) %>% drop_na() %>% 
  filter(oncotree_code %in% cohort_table_1$ONCOTREE_CODE) %>%
  group_by(oncotree_code) %>% mutate(n_type = n_distinct(sample_id)) %>%
  group_by(Hugo_Symbol, oncotree_code, biallelic_label) %>% add_count() %>% mutate(n_events = sum(os_status)) %>%
  group_by(Hugo_Symbol, oncotree_code) %>% filter(n_type >= 100 & n_distinct(biallelic_label) == 3 & min(n) >= 10 & min(n_events) >= 5) %>% ungroup() %>%
  select(Hugo_Symbol, oncotree_code) %>% unique() %>% as.data.table()

n_distinct(pairs_to_test$biallelic_label$Hugo_Symbol)
n_distinct(pairs_to_test$biallelic_label$oncotree_code)
nrow(pairs_to_test$biallelic_label)


```

```{r}

plot(eulerr::euler(list(mutation_class = paste0(pairs_to_test$mutation_class$Hugo_Symbol, "_", pairs_to_test$mutation_class$oncotree_code),
                        biallelic_label = paste0(pairs_to_test$biallelic_label$Hugo_Symbol, "_", pairs_to_test$biallelic_label$oncotree_code))), quantities = T)

```

## OS by \_ function

For each gene/cancer type pair, return:

-   survfit object (KM)

-   coxph object

    -   OS \~ [variable] + disease status + sex + age at dx + level 1 actionable alt + FGA + TMB + MSI (multivariate)

-   tidy coxph object with ns

```{r}

os_by_var <- function(combo = NULL, gene_name, cancer_type_name, variable, ref = "wt", covars = c("disease_status", "sex", "age_at_dx", "oncokb_level_1", "fga", "tmb", "msi_score"), data = data_os_zyg) {
  
  if(!is.null(combo)) {
    gene_name <- combo[1]
    cancer_type_name <- combo[2]
  }
  
  d <- data_os_zyg %>%
    filter(oncotree_code == cancer_type_name & Hugo_Symbol == gene_name) %>%
    select(os_months, os_status, all_of(c(variable, covars))) %>% 
    mutate(across(where(is.character), ~ fct_drop(fct_infreq(.x)))) %>%
    mutate(disease_status = fct_relevel(disease_status, "Primary")) %>%
    drop_na() %>% 
    mutate(!!as.name(variable) := fct_relevel(!!as.name(variable), ref))
  
  # remove covars with only one level
  to_remove <- d %>% select(where(~all(n_distinct(.)==1))) %>% names()
  d <- d %>% select(-all_of(to_remove))
  
  covars <- covars[covars %in% names(d)]
  
  # consolidate small levels
  d <- d %>% mutate(across(where(is.factor) & !variable, ~ drop_small_factor_levels_col(.x, min_n = 10, verbose = F)))
  
  counts_t <- d %>% group_by(!!as.name(variable)) %>% count() %>% pivot_wider(names_from = !!as.name(variable), names_prefix = "n_", values_from = n)

  surv <- survfit2(as.formula(paste0("Surv(os_months, os_status) ~ ", variable)), data = d)
  #surv
  #surv %>% ggsurvplot()

  model <- coxph(as.formula(paste0("Surv(os_months, os_status) ~ ", paste0(variable, " + ", paste0(covars, collapse = " + ")))), data = d)
  #ggforest(model)
  
  results_table <- tidy(model, exponentiate = T, conf.int = T) %>%
    mutate(gene = gene_name, cancer_type = cancer_type_name, ref = ref) %>%
    cbind(counts_t) %>%
    select(gene, cancer_type, ref, term, everything())

  return(list(uv_model = surv, mv_model = model, input_data = d, res_table = results_table))
  
  }

```

```{r}

# keap1 luad
res <- os_by_var(gene_name = "KEAP1", cancer_type_name = "LUAD", variable = "biallelic_label")
ggforest(res$mv_model, data = res$input_data)
res$res_table
save(res, file = paste0(outdir, "os_zyg_screen_keap1_luad.Robj"))

res <- os_by_var(gene_name = "KEAP1", cancer_type_name = "LUAD", variable = "mutation_class")
ggforest(res$mv_model, data = res$input_data)
res$res_table
save(res, file = paste0(outdir, "os_mutclass_screen_keap1_luad.Robj"))


```

# OS by mutation class

## WT ref

```{r}

gene_type_list <- as.list(as.data.frame(t(pairs_to_test$mutation_class)))
os_by_mut_class <- mclapply(gene_type_list, os_by_var, variable = "mutation_class", mc.cores = 6)

```

```{r}

mut_class_names <- unlist(lapply(gene_type_list, function(x) {paste0(x[1], "_", x[2])}))
names(mut_class_names) <- NA
names(os_by_mut_class) <- mut_class_names

```

```{r}

os_by_mut_class_table <- do.call(rbind, lapply(os_by_mut_class, function(x) {return(x$res_table)}))
os_by_mut_class_table <- os_by_mut_class_table %>% 
  remove_rownames() %>%
  group_by(gene, cancer_type) %>%
  mutate(p.adj = p.adjust(p.value, "fdr", n = length(gene_type_list))) %>% ungroup() %>%
  mutate(p.adj.signif.label = case_when(p.adj < 0.05 ~ "p.adj < 0.05",
                                        p.value < 0.05 ~ "p < 0.05", 
                                        TRUE ~ "ns"))

os_by_mut_class_table %>% filter(grepl("mutation_classvus", term)) %>% arrange(p.value)

fwrite(os_by_mut_class_table, paste0(outdir, "os_by_mut_class.csv"))

```

## VUS ref

```{r}

os_by_mut_class2 <- mclapply(gene_type_list, os_by_var, variable = "mutation_class", ref = "vus", mc.cores = 6)
names(os_by_mut_class2) <- mut_class_names

os_by_mut_class2$KEAP1_LUAD

```

```{r}

os_by_mut_class_table2 <- do.call(rbind, lapply(os_by_mut_class2, function(x) {return(x$res_table)}))
os_by_mut_class_table2 <- os_by_mut_class_table2 %>% 
  remove_rownames() %>%
  group_by(gene, cancer_type) %>%
  mutate(p.adj = p.adjust(p.value, "fdr", n = length(gene_type_list))) %>% ungroup() %>%
  mutate(p.adj.signif.label = case_when(p.adj < 0.05 ~ "p.adj < 0.05",
                                        p.value < 0.05 ~ "p < 0.05", 
                                        TRUE ~ "ns"))

os_by_mut_class_table2 %>% filter(grepl("mutation_classdriver", term)) %>% arrange(p.value)

fwrite(os_by_mut_class_table2, paste0(outdir, "os_by_mut_class_vus_ref.csv"))

```

## Combined (mutated)

```{r}

gene_type_list <- as.list(as.data.frame(t(pairs_to_test$mutation_class)))
os_by_mut_class_alt <- mclapply(gene_type_list, os_by_var, variable = "mutation_class_combined", mc.cores = 6)

```

```{r}

names(os_by_mut_class_alt) <- mut_class_names

```

```{r}

os_by_mut_class_alt_table <- do.call(rbind, lapply(os_by_mut_class_alt, function(x) {return(x$res_table)}))
os_by_mut_class_alt_table <- os_by_mut_class_alt_table %>% 
  remove_rownames() %>%
  group_by(gene, cancer_type) %>%
  mutate(p.adj = p.adjust(p.value, "fdr", n = length(gene_type_list))) %>% ungroup() %>%
  mutate(p.adj.signif.label = case_when(p.adj < 0.05 ~ "p.adj < 0.05",
                                        p.value < 0.05 ~ "p < 0.05", 
                                        TRUE ~ "ns"))

os_by_mut_class_alt_table

fwrite(os_by_mut_class_alt_table, paste0(outdir, "os_by_mut_class_alt.csv"))

```

## 

# OS by zygosity

## WT ref

```{r}

gene_type_list <- as.list(as.data.frame(t(pairs_to_test$biallelic_label)))
os_by_zyg <- mclapply(gene_type_list, os_by_var, variable = "biallelic_label", mc.cores = 6)

```

```{r}

zyg_names <- unlist(lapply(gene_type_list, function(x) {paste0(x[1], "_", x[2])}))
names(zyg_names) <- NA
names(os_by_zyg) <- zyg_names

```

```{r}

os_by_zyg_table <- do.call(rbind, lapply(os_by_zyg, function(x) {return(x$res_table)}))
os_by_zyg_table <- os_by_zyg_table %>% 
  remove_rownames() %>%
  group_by(gene, cancer_type) %>%
  mutate(p.adj = p.adjust(p.value, "fdr", n = length(gene_type_list))) %>% ungroup() %>%
  mutate(p.adj.signif.label = case_when(p.adj < 0.05 ~ "p.adj < 0.05",
                                        p.value < 0.05 ~ "p < 0.05", 
                                        TRUE ~ "ns"))

os_by_zyg_table %>% filter(grepl("biallelic_labelbiallelic", term)) %>% arrange(p.value)

fwrite(os_by_zyg_table, paste0(outdir, "os_by_zyg.csv"))

```

## monoallelic ref

```{r}

os_by_zyg2 <- mclapply(gene_type_list, os_by_var, variable = "biallelic_label", ref = "monoallelic", mc.cores = 6)
names(os_by_zyg2) <- zyg_names

```

```{r}

os_by_zyg_table2 <- do.call(rbind, lapply(os_by_zyg2, function(x) {return(x$res_table)}))
os_by_zyg_table2 <- os_by_zyg_table2 %>% 
  remove_rownames() %>%
  group_by(gene, cancer_type) %>%
  mutate(p.adj = p.adjust(p.value, "fdr", n = length(gene_type_list))) %>% ungroup() %>%
  mutate(p.adj.signif.label = case_when(p.adj < 0.05 ~ "p.adj < 0.05",
                                        p.value < 0.05 ~ "p < 0.05", 
                                        TRUE ~ "ns"))

os_by_zyg_table2 %>% filter(grepl("biallelic_labelbiallelic", term)) %>% arrange(p.value)

fwrite(os_by_zyg_table2, paste0(outdir, "os_by_zyg_monoallelic_ref.csv"))

```

## Combined (altered)

```{r}

gene_type_list <- as.list(as.data.frame(t(pairs_to_test$biallelic_label)))
os_by_zyg_alt <- mclapply(gene_type_list, os_by_var, variable = "biallelic_label_combined", mc.cores = 6)

```

```{r}

names(os_by_zyg_alt) <- zyg_names

```

```{r}

os_by_zyg_alt_table <- do.call(rbind, lapply(os_by_zyg_alt, function(x) {return(x$res_table)}))
os_by_zyg_alt_table <- os_by_zyg_alt_table %>% 
  remove_rownames() %>%
  group_by(gene, cancer_type) %>%
  mutate(p.adj = p.adjust(p.value, "fdr", n = length(gene_type_list))) %>% ungroup() %>%
  mutate(p.adj.signif.label = case_when(p.adj < 0.05 ~ "p.adj < 0.05",
                                        p.value < 0.05 ~ "p < 0.05", 
                                        TRUE ~ "ns"))

os_by_zyg_alt_table

fwrite(os_by_zyg_alt_table, paste0(outdir, "os_by_zyg_alt.csv"))

```
